Single Site, Single Disturbance

### 2020-10-03 SingleSite_SingleDisturb.Rmd
Added ridgeline and violin visualizations of multimodal Analysis

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

@@

```{r}
FilterSSSD <- "
model{
  N ~ dnorm(muIC,paIC)  ## initial conditions
  ## Forecast
  muN <- N + R*N*(1-N/K)
  muD ~ dnorm(mu0,pa0)
  D ~ dbern(p)
  mu <- D*muD + (1-D)*muN
  n ~ dnorm(mu,q)
  ## Analysis
  Y ~ dnorm(n,r)
}"
## Forecast Cycle
CI.fN = matrix(NA,5,NT)
CI.fD = matrix(NA,5,NT)
Xl = nIC         ## set IC
pl = 1/n.sd^2
Dbar = 0
#mu0 = data$mu0   ## set disturbed state
#pa0 = data$pa0
#aqp = data$aq
#bqp = data$bq
#arp = data$ar
#brp = data$br
for(t in 1:NT){
  ## Forecast + Analysis
  update = list(Y=data$Y[t],
                muIC=Xl,paIC=pl,
                mu0=data$mu0,pa0=data$pa0,
                #aq=aqp[t],bq=bqp[t],ar=arp[t],br=brp[t],
                R=r, K=K,p = p.dist, q = 1/n.sd^2, r=1/obs.sd^2     ## hard coding these to start
                )
  mod <- jags.model(file=textConnection(FilterSSSD),
                   data=update,
                  n.adapt=1000,n.chains=3)#,
#                  init=list(X.ic=Xlf))
 jdat <- coda.samples(mod,variable.names=c("n","D","muN","muD"),n.iter=10000) 
# plot(jdat)

@@

#  mq = mean(dat[,"q"])
#  vq = var(dat[,"q"])
#  aqp[t+1] = mq*mq/vq
#  bqp[t+1] = mq/vq
#  mr = mean(dat[,"r"])
#  vr = var(dat[,"r"])
#  arp[t+1] = mr*mr/vr
#  brp[t+1] = mr/vr
}

@@

```{r}
## comparison to EnKF
plot(data$Y,ylim = range(rbind(CIN,data$Y)),xlim=xlim,type='n',
     ylab="LAI",xlab="time")
polygon(cbind(c(time, rev(time), 1), c(an0.stats[2, ], rev(an0.stats[4, ]),
  an0.stats[1, 1])), border = NA, col = '#FFBF6C')
polygon(cbind(c(time, rev(time), 1), c(an.stats[2, ], rev(an.stats[4, ]),
  an.stats[1, 1])), border = NA, col = ecoforecastR::col.alpha('#b2df8a',0.7))
#ecoforecastR::ciEnvelope(1:NT,CIN[1,],CIN[3,],col=rgb(col[1],col[2],col[3],0.3*256,maxColorValue=256))
ecoforecastR::ciEnvelope(1:NT,CI.fN[2,],CI.fN[4,],col=rgb(col[1],col[2],col[3],0.3*256,maxColorValue=256))

@@

## Todo
-   Visualize bimodal forecast and analysis distributions through time for simple example
-   Tweak model to have a simple C budget -- where does the C go when the model doesn't assimilate disturbance?
-   Apply to SIPNET SDA

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rjags)
library(ecoforecastR)
library(tidyverse)
```

@@

```{r}
FilterSSSD <- "
model{
  N ~ dnorm(muIC,paIC)  ## initial conditions
  ## Forecast
  muN <- N + R*N*(1-N/K)
  muD ~ dnorm(mu0,pa0)
  D ~ dbern(p)
  mu <- D*muD + (1-D)*muN
  n ~ dnorm(mu,q)
  ## Analysis
  Y ~ dnorm(n,r)
}"
## Forecast Cycle
CI.fN = matrix(NA,5,NT)
CI.fD = matrix(NA,5,NT)
Xl = nIC         ## set IC
pl = 1/n.sd^2
Dbar = 0
#mu0 = data$mu0   ## set disturbed state
#pa0 = data$pa0
#aqp = data$aq
#bqp = data$bq
#arp = data$ar
#brp = data$br
Analysis <- list()
for(t in 1:NT){
  ## Forecast + Analysis
  update = list(Y=data$Y[t],
                muIC=Xl,paIC=pl,
                mu0=data$mu0,pa0=data$pa0,
                #aq=aqp[t],bq=bqp[t],ar=arp[t],br=brp[t],
                R=r, K=K,p = p.dist, q = 1/n.sd^2, r=1/obs.sd^2     ## hard coding these to start
                )
  mod <- jags.model(file=textConnection(FilterSSSD),
                   data=update,
                  n.adapt=1000,n.chains=3)#,
#                  init=list(X.ic=Xlf))
 jdat <- coda.samples(mod,variable.names=c("n","D","muN","muD"),n.iter=10000) 
# plot(jdat)

@@

#  mq = mean(dat[,"q"])
#  vq = var(dat[,"q"])
#  aqp[t+1] = mq*mq/vq
#  bqp[t+1] = mq/vq
#  mr = mean(dat[,"r"])
#  vr = var(dat[,"r"])
#  arp[t+1] = mr*mr/vr
#  brp[t+1] = mr/vr

  ## save Analysis posterior
  Analysis[[t]] <- dat

@@

### Comparision to EnKF
```{r}
plot(data$Y,ylim = range(rbind(CIN,data$Y)),xlim=xlim,type='n',
     ylab="LAI",xlab="time")
polygon(cbind(c(time, rev(time), 1), c(an0.stats[2, ], rev(an0.stats[4, ]),
  an0.stats[1, 1])), border = NA, col = '#FFBF6C')
polygon(cbind(c(time, rev(time), 1), c(an.stats[2, ], rev(an.stats[4, ]),
  an.stats[1, 1])), border = NA, col = ecoforecastR::col.alpha('#b2df8a',0.7))
#ecoforecastR::ciEnvelope(1:NT,CIN[1,],CIN[3,],col=rgb(col[1],col[2],col[3],0.3*256,maxColorValue=256))
ecoforecastR::ciEnvelope(1:NT,CI.fN[2,],CI.fN[4,],col=rgb(col[1],col[2],col[3],0.3*256,maxColorValue=256))

@@

### Ridgeline plot of Analysis
```{r}
## organize Analysis state var (n) into a wide format
n.wide <- sapply(Analysis,function(x){x[,"n"]})
colnames(n.wide) <- paste0("t",1:NT)
n.wide <- as.data.frame(n.wide)

## transform to a long format
n.long <- tidyr::pivot_longer(n.wide[,xlim[1]:xlim[2]],
                              cols=1:(diff(xlim)+1),
                              names_to = "time",names_prefix = "t",
                              values_to = "n")
n.long$time <- as.numeric(n.long$time)

## ridge plot
ggplot(n.long) + aes(x=n,y=time,group=time) +
  ggridges::geom_density_ridges(bandwidth = 0.02) +
  labs(x = "LAI", y = "Time")

## violin plot
n.long %>%
 filter(n >= 0 & n <= 8) %>%
 ggplot() +
 aes(x = "", y = n, group = time) +
 geom_violin(adjust = 0.5, scale = "area", fill = "#0c4c8a") +
 labs(x = "Time", y = "LAI") +
 theme_minimal()

## Todo
-   Tweak model to have a simple C budget -- where does the C go when the model doesn't assimilate disturbance?
-   Apply to SIPNET SDA
